---
title: "Tech 1. Data Prep."
output:
  html_document: default
  word_document: default
  pdf_document: default
---

This workflow file takes the source data, recodes variables and prepares descriptive tables for comparison. 

The original Tables 1-3 were extracted from the original study PDF using Adobe Acrobat 'copy with formatting' function and placed into the document 'orig_tables.xlsx'.

Dependent Variables

V272: Compared with five years ago, has your household's standard of living fallen a great deal, fallen a little, stayed about the same, risen a little, or has it risen a lot?

fallen a great deal	1
fallen a little	2
stayed about the same	3
risen a little	4
risen a lot	5
(Don't know)	8

V273: And looking ahead over the next five years, do you think that your household's standard of living will fall a great deal from its current level, fall a little, stay about the same as it is now, rise a little, or rise a lot from its current level?

will fall a great deal	1
will fall a little	2
will stay about the same	3
will rise a little	4
will rise a lot	5
(Don't know)	8


```{r setup}
pacman::p_load("tidyverse","readxl","foreign","skimr","fastDummies","kableExtra","MASS","sjPlot","webshot","knitr")
```

## Import Data

Datafile acquired after emails to Horvat then to Evans then to the ultimate source of this aggregate file "Mass_Public_Surveys_1993-2007.dta" with Ksenia Northmore-Ball (Nov. 8th, 2020)

```{r import, warning = F, message = F}
# original data
df_origa <- readstata13::read.dta13("data/Mass_Public_Surveys_1993-2007.dta")
df_orig_labels <- readstata13::read.dta13("data/Mass_Public_Surveys_1993-2007.dta", generate.factors = T)

# original tables 1 & 2 from the article
Tbl1_orig <- read_xlsx("data/orig_tables.xlsx", sheet = "Tbl1", col_names = F)
Tbl2_orig <- read_xlsx("data/orig_tables.xlsx", sheet = "Tbl2", col_names = F)
Tbl3_orig <- read_xlsx("data/orig_tables.xlsx", sheet = "Tbl3", col_names = F)

# codebook
codebook <- read.csv("data/codebook.csv", header = T)
```

|v270 Main source of income| Code|
|---|-|
|Earnings from employment (own or partner's)	|1|
|Pensions and benefits	|2|
|Student stipend	|3|
|Other state benefit	|4|
|Interest from savings or property	|5|
|Dependent on family/relatives	|6|

|v200 Work Status| Code|
|---|-|
|in paid work (including self-employment)	|1|
|full-time student 	|2|
|in military service	|3|
|unemployed	|4|
|permanently sick or disabled	|5|
|completely retired from work	|6|
|looking after the home	|7|
|other	|8|
|NA	|9|
|demobilized	|10|
|vacation without salary	|11|

Education = "Education is measured by three categories: low,
middle, and higher education. Low education means
no educational qualifications beyond the compulsory
level. Middle education corresponds to completed
secondary education and higher education corresponds
to completed further or university education." (p. 713)

ISCED-97
0 pre-primary
1 primary/1st stage basic ed
2 lower secondary
3 upper secondary
4 post-secondary non-tertiary
5 1st stage tertiary
6 2nd stage tert

**v269 income per month**
Income recoded into terciles by country-year (see footnote 3)

**Class**

"Social class is measured by a six-category version
of the Erikson–Goldthorpe class schema (Erikson and
Goldthorpe, 1992), based on occupational measures of
class position: service class, routine non-manual workers,
self-employed workers, manual supervisors and
skilled manual workers, semi-skilled and unskilled
manual workers, and farmers or agricultural workers.
A residual category ‘never had a paying job’ denotes
respondents whose social class was ambiguous or
missing but who reported never having been in paid
employment elsewhere in the survey. Women with
missing social class data were classified according to
their husband’s class. Previous research on Eastern
Europe suggests that occupational measures of class
position perform adequately in the Eastern European
context and successfully differentiate individuals in
terms of their level of income, their degree of economic
security, and chances of economic advancement
(Evans, 1997; Evans and Mills, 1999)."

Students dropped

**Age**
cat_age: cross-temporally and cross-nationally consistent age categories: 
1:  -29; 2: 30-44; 3: 45-59; 4: 60+


## Recode Analysis Variables

```{r recode, warning = F}
df_orig <- df_origa %>%
  mutate(stdliv_past5 = car::recode(v272, "8 = 3"), # don't knows were recoded into middle cat in original study
         stdliv_past5_di = car::recode(stdliv_past5, "c(1,2) = 1; c(4,5) = 2; c(3) = 3"),
         stdliv_next5 = car::recode(v273, "8 = 3"),
         stdliv_next5_di = car::recode(stdliv_next5, "c(1,2) = 1; c(4,5) = 2; c(3) = 3"),
         noway_future_improve = as.numeric(as.character(ifelse(v276 == "no any way", 1, 0))),
         mkt_econ_eval = car::recode(v4, "'dont know' = 3"),
         mkt_econ_eval_di = car::recode(mkt_econ_eval, "c('very positively','positively') = 1; c('negatively','very negatively') = 2; c('neither positively nor negatively') = 3"),
         student = ifelse(v200 == 2 | v200 == 9, NA, 0), # remove students and NA's
         wave = as.numeric(car::recode(year, "c('1993','1994','1995','1996') = '1993'; c('2007') = '2007'; c('1997','1998','2001','2002','2003','2004','2005','2006') = NA")),
         female = as.numeric(v298) - 1,
         pensions = ifelse(is.na(v270), NA, ifelse(v270==2, 1, 0)),
         unemployed = ifelse(v200 == 8, NA, ifelse(v200==4, 1, 0)),
         car_owner = ifelse(v262 ==1, 1, 0),
         education = car::recode(std_education, "c(0,1,2) = 1; c(3,4) = 2; c(5,6) = 3; c(99) = NA"), # some had 99 in the std_education variable still, made into primary
         education_a = car::recode(std_education, "c(1,2,95,96) = 1; c(3,4,5,6,8) = 2; c(7,9,10,11,12) = 3; c(98,99,14) = NA"), # 4 and 5 are questionable categories here
         education = ifelse(cntry == "russia" & year == 1996, education_a, education),
         # There is a problem with Russia in 1996, seems that it was not recoded into ISCED
         EGP6 = car::recode(rclass10, "'Missing in 93-03 data' = 8;'semi-unskilld manual' = 5; 'skilled manual' = 4; 'higher controllers' = 1;'lo controllers' = 1;  'routine nonmanual'= 2; 'sempl without empl' = 3; 'sempl with emp' = 4; 'selfempl farm' = 6; 'farm labor'=6; 'manual supervisor' = 4; 'Missing Occupation Code 07' = 8; 'Not ISKO codes' = 8; 'ISKO Coded - no match' = 8; 'Never had a paying job' = 7"),
         EGP6 = ifelse(EGP6 == 8 & r2class10 == "never had a paying job", "7", EGP6),
         EGP6 = ifelse(is.na(v201), EGP6, ifelse(v201 == "no", "7", EGP6)), # never had a paid job
         EGP6 = ifelse(EGP6 == 8, NA, EGP6)
         ) %>%
  group_by(cntry, year) %>%
  mutate(income = ifelse(is.na(v269), 4, ntile(v269, 3))) %>%
  ungroup()
```

## Weights

There are some curiosities with the weights. There are no weights for 1993 and other weights are NA meaning that full weighting is not possible. I recode NA to 1 for now to simply preserve cases.

```{r weights, warning = F}
# some weights are NA, replace with 1 (actually there are no weights for 1993, so this is not really helpful)
df_orig$weight <- ifelse(is.na(df_orig$v461), 1, df_orig$v461)
```

## Complete Dataframe

And generate group descriptive means

```{r descgrp, warning = F}
# create complete cases df
df_orig_complete <- df_orig[!is.na(df_orig["education"]),]
# df_orig_complete <- df_orig_complete[!is.na(df_orig_complete["income"]),]
df_orig_complete <- df_orig_complete[!is.na(df_orig_complete["EGP6"]),]
df_orig_complete <- df_orig_complete[!is.na(df_orig_complete["female"]),]
df_orig_complete <- df_orig_complete[!is.na(df_orig_complete["wave"]),]
df_orig_complete <- df_orig_complete[!is.na(df_orig_complete["student"]),]
df_orig_complete <- df_orig_complete[!is.na(df_orig_complete["pensions"]),]
df_orig_complete <- df_orig_complete[!is.na(df_orig_complete["unemployed"]),]
df_orig_complete <- df_orig_complete[!is.na(df_orig_complete["car_owner"]),]
df_orig_complete <- df_orig_complete[!is.na(df_orig_complete["stdliv_past5"]),]
df_orig_complete <- df_orig_complete[!is.na(df_orig_complete["stdliv_next5"]),]
df_orig_complete <- df_orig_complete[!is.na(df_orig_complete["mkt_econ_eval"]),]
df_orig_complete <- df_orig_complete[!is.na(df_orig_complete["noway_future_improve"]),]
# create factor dummies
df_orig_complete <- dummy_cols(df_orig_complete, select_columns = c("EGP6", "income", "education","stdliv_past5_di","stdliv_next5_di","mkt_econ_eval_di"))

#create new group ID
df_orig_complete$group <- df_orig_complete$cat_age + (100*df_orig_complete$wave)

# cases per group

cases <- df_orig_complete %>%
  group_by(group) %>%
  count() %>%
  ungroup() %>%
  dplyr::select(-group) %>%
  t()

  #df_orig <- select(df_orig, v201, rclass10, EGP6, everything())
```

## Replicate Table 1

```{r Tbl1_rep}
# get weighted means by group
Tbl1_rep <- apply(df_orig_complete[,c("education_1","education_2","education_3","income_1","income_2","income_3","EGP6_1","EGP6_2","EGP6_3","EGP6_4","EGP6_5","EGP6_6","EGP6_7","female")], 2, function(x) {sapply(split(data.frame(df_orig_complete[,"weight"], x), df_orig_complete$group), function(y) weighted.mean(y[,2], w = y$weight))})

Tbl1_rep <- as.data.frame(t(round(Tbl1_rep*100, 0)))

Tbl1_rep_unw <- apply(df_orig_complete[,c("education_1","education_2","education_3","income_1","income_2","income_3","EGP6_1","EGP6_2","EGP6_3","EGP6_4","EGP6_5","EGP6_6","EGP6_7","female")], 2, function(x) {sapply(split(data.frame(df_orig_complete[,"weight"], x), df_orig_complete$group), function(y) mean(y[,2], w = y$weight))})

Tbl1_rep_unw <- as.data.frame(t(round(Tbl1_rep_unw*100, 0)))

Tbl1_rep[nrow(Tbl1_rep)+1,] <- cases

Tbl1_rep_unw[nrow(Tbl1_rep_unw)+1,] <- cases

rownames(Tbl1_rep)[15] <- "N"
rownames(Tbl1_rep_unw)[15] <- "N"

# create csv for easy importing in final report
write.csv(Tbl1_rep, here::here("results", "Tbl1_rep_unw.csv"), row.names = F)

Tbl1_rep_out <- kable_styling(kable(Tbl1_rep, col.names = c("Wave 1, Age 18-29","Wave 1, Age 30-49","Wave 1, Age 50-59","Wave 1, Age 60+","Wave 2, Age 18-29","Wave 2, Age 30-49","Wave 2, Age 50-59","Wave 2, Age 60+"), caption = "Table 1R. Replicated Weighted. Age and socio-demographic outcomes weighted in percentages by age group"))
save_kable(Tbl1_rep_out, file = "results/Tbl1_rep.htm")

Tbl1_rep_out_unw <- kable_styling(kable(Tbl1_rep, col.names = c("Wave 1, Age 18-29","Wave 1, Age 30-49","Wave 1, Age 50-59","Wave 1, Age 60+","Wave 2, Age 18-29","Wave 2, Age 30-49","Wave 2, Age 50-59","Wave 2, Age 60+"), caption = "Table 1R. Replicated Unweighted. Age and socio-demographic outcomes weighted in percentages by age group"))


#webshot("results/Tbl1_rep.htm", file = "results/Tbl1_rep.png")
```

### Table 1R Weighted
```{r Tbl1_repo}
Tbl1_rep_out

```
### Table 1R Unweighted

```{r Tbl1_repounw}
Tbl1_rep_out_unw
```

## Table 1 Original

```{r tbl1_orig_comp}
Tbl1_orig <- Tbl1_orig[-c(1:3),]
colnames(Tbl1_orig) <- c("Variable","y93-96 a18-29","y93-96 a30-44","y93-96 a 45-59","y93-96 a60+","y07 18-29","y07 30-44","y07 45-59","y07 60+")

Tbl1_orig[,2:9] <- lapply(Tbl1_orig[,2:9], function (x) ifelse(is.na(x), "", x))

Tbl2_orig <- Tbl2_orig[-c(1:2),]

colnames(Tbl2_orig) <- c("Variable","y93-96 a18-29","y93-96 a30-44","y93-96 a 45-59","y93-96 a60+","y07 18-29","y07 30-44","y07 45-59","y07 60+")

Tbl2_orig[,2:9] <- lapply(Tbl2_orig[,2:9], function (x) ifelse(is.na(x), "", x))

Tbl1_orig$Variable <- c(NA, "Education_1_Low", "Education_2_Mid", "Education_3_Hi", NA, "Income_1_Low", "Income_2_Mid", "Income_3_Hi", NA, "EGP6_1_Service", "EGP6_2_Rtn_NonMan", "EGP6_3_SelfEmp", "EGP6_4_SkilMan", "EGP6_5_Unskilled", "EGP6_6_Farmers", "EGP6_7_NeverHadJob", "Female", "N")

Tbl1_orig <- subset(Tbl1_orig, !is.na(Variable))

kable_styling(kable(Tbl1_orig, caption = "Table 1O. Original. Age and socio-demographic outcomes weighted in percentages by age group"))
```

The results do not match very well, especially in the education category. Actually it appears that only 2007 has weights. Running Table 1 without weights changes very little. Therefore, a proper test requires hand coding by country; however, the country-specific education codes in the codebook for v198 are incorrect, e.g., I tested Bulgaria and Belarus 1993. Belarus looks ok except two missing codes are not present, but Bulgaria has several extra codes than what is listed in the codebook provided (values 8-12). Therefore, this hand coding is not possible. Russia has different codes altogether, therefore I had to guess how to code it under the assumption the variable was similarly ordinal.

## Education Peculiarities

### ISCED in all countries except Russia

```{r isced check}
unique(df_orig$std_education[df_orig$cntry != "russia"])
```

### Russia not in ISCED

Extra categories not in ISCED-97 or in other countries

```{r russiaisced}
unique(df_orig$std_education[df_orig$cntry == "russia"])
```
### Check original education variable

#### Russia

Not coded following original codebook (see "data" folder). At least one category is present in data ("14") that is not in codebook

RUSSIA 1993	
V198  highest education level
1  primary school
2  secondary school
3  high school
4  professional courses
5  vocational school
6  technical secondary school
7  vocational post-school
8  technical college
9  incomplete high education(institute,university,academy)
10  high education(institute,university,academy)
11  additional training courses
12  degree
95  no educational qualifications
96  never went to school
98  don't know


```{r russia}
unique(df_orig$v198[df_orig$cntry == "russia"])
```

#### Belarus

BELARUS	 1993
V198  education
1  elementary school
2  junior high school
3  high school
4  professional training courses
5  regular factory-and-workshop school,industrial training scho
6  industrial training high school
7  college (for nurses, elementary school teachers, musicians)
8  technical college
9  bachelor's degree
10  master's degree (university, academy etc)
11  postgraduate courses
12  Ph. Degree
95  no certificates of education of any kind
96  never went to school
98  don't know


```{r belarus}
unique(df_orig$v198[df_orig$cntry == "belarus"])
```

#### Bulgaria 

BULGARIA 1993
V198  education
1  NO COMPLETED ED.
2  ELEMENTARY
3  PRIMARY
4  HIGH SCHOOL
5  SECONDARY VOCATIONAL
6  COLLEGE
7  HIGHER
98  DON T KNOW

```{r bulgaria}
unique(df_orig$v198[df_orig$cntry == "belarus"])
```

## Table 2 Replicated

```{r tbl2_rep}
# get weighted means by group
Tbl2_rep <- apply(df_orig_complete[,c("pensions","unemployed","car_owner")], 2, function(x) {sapply(split(data.frame(df_orig_complete[,"weight"], x), df_orig_complete$group), function(y) weighted.mean(y[,2], w = y$weight))})

Tbl2_rep <- as.data.frame(t(round(Tbl2_rep, 3)))

Tbl2_rep[nrow(Tbl2_rep)+1,] <- cases

rownames(Tbl2_rep)[4] <- "N"

# create csv for easy importing in final report
write.csv(Tbl2_rep, here::here("results", "Tbl2_rep.csv"), row.names = F)

Tbl2_rep[1,] <- round(as.numeric(Tbl2_rep[1,])*100,0)
Tbl2_rep[2,] <- round(as.numeric(Tbl2_rep[2,])*100,0)
Tbl2_rep[3,] <- round(as.numeric(Tbl2_rep[3,])*100,0)
Tbl2_rep[4,] <- round(as.numeric(Tbl2_rep[4,]),0)

kable_styling(kable(Tbl2_rep, caption = "Table 2R. Replicated. Age-based inequality in resources in percentage", col.names = c("Wave 1, Age 18-29", "Wave 1, Age 30-44", "Wave 1, Age 45-59", "Wave 1, 60+", "Wave 2, Age 18-29", "Wave 2, Age 30-44", "Wave 2, Age 45-59", "Wave 2, 60+")))
```

## Table 2. Original

```{r Tbl2}
kable_styling(kable(Tbl2_orig, caption = "Table 2O. Original. Age-based inequality in resources in percentage"))
```
## Table 3R. Replicated Weights.

```{r tbl3_rep}
Tbl3_rep <- apply(df_orig_complete[,c("stdliv_past5_di_1","stdliv_past5_di_2","stdliv_next5_di_1","stdliv_next5_di_2","noway_future_improve","mkt_econ_eval_di_1","mkt_econ_eval_di_2")], 2, function(x) {sapply(split(data.frame(df_orig_complete[,"weight"], x), df_orig_complete$group), function(y) weighted.mean(y[,2], w = y$weight))})

Tbl3_rep <- as.data.frame(t(round(Tbl3_rep, 3)))

Tbl3_rep[nrow(Tbl3_rep)+1,] <- cases

rownames(Tbl3_rep)[8] <- "N"

Tbl3_rep[1,] <- round(as.numeric(Tbl3_rep[1,])*100,0)
Tbl3_rep[2,] <- round(as.numeric(Tbl3_rep[2,])*100,0)
Tbl3_rep[3,] <- round(as.numeric(Tbl3_rep[3,])*100,0)
Tbl3_rep[4,] <- round(as.numeric(Tbl3_rep[4,])*100,0)
Tbl3_rep[5,] <- round(as.numeric(Tbl3_rep[5,])*100,0)
Tbl3_rep[6,] <- round(as.numeric(Tbl3_rep[6,])*100,0)
Tbl3_rep[7,] <- round(as.numeric(Tbl3_rep[7,])*100,0)
Tbl3_rep[8,] <- round(as.numeric(Tbl3_rep[8,]),0)

kable_styling(kable(Tbl3_rep, caption = "Table 3R. Replicated Weights. Age-based inequality in economic experience in percentages", col.names = c("Wave 1, Age 18-29", "Wave 1, Age 30-44", "Wave 1, Age 45-59", "Wave 1, 60+", "Wave 2, Age 18-29", "Wave 2, Age 30-44", "Wave 2, Age 45-59", "Wave 2, 60+")))
```

Here I can reproduce Table 3 almost exactly, but only when I **do not use weights**, therefore, it is unclear what it means in the text when they claim that the descriptives are weighted.

```{r tbl3_rep_no}
Tbl3_rep_n <- apply(df_orig_complete[,c("stdliv_past5_di_1","stdliv_past5_di_2","stdliv_next5_di_1","stdliv_next5_di_2","noway_future_improve","mkt_econ_eval_di_1","mkt_econ_eval_di_2")], 2, function(x) {sapply(split(data.frame(df_orig_complete[,"weight"], x), df_orig_complete$group), function(y) mean(y[,2], w = y$weight))})

Tbl3_rep_n <- as.data.frame(t(round(Tbl3_rep_n, 3)))

Tbl3_rep_n[nrow(Tbl3_rep_n)+1,] <- cases

rownames(Tbl3_rep_n)[7] <- "N"

Tbl3_rep_n[1,] <- round(as.numeric(Tbl3_rep_n[1,])*100,0)
Tbl3_rep_n[2,] <- round(as.numeric(Tbl3_rep_n[2,])*100,0)
Tbl3_rep_n[3,] <- round(as.numeric(Tbl3_rep_n[3,])*100,0)
Tbl3_rep_n[4,] <- round(as.numeric(Tbl3_rep_n[4,])*100,0)
Tbl3_rep_n[5,] <- round(as.numeric(Tbl3_rep_n[5,])*100,0)
Tbl3_rep_n[6,] <- round(as.numeric(Tbl3_rep_n[6,])*100,0)
Tbl3_rep_n[7,] <- round(as.numeric(Tbl3_rep_n[7,])*100,0)
Tbl3_rep_n[8,] <- round(as.numeric(Tbl3_rep_n[8,]),0)

# create csv for easy importing in final report
write.csv(Tbl3_rep_n, here::here("results", "Tbl3_rep_n.csv"), row.names = F)

kable_styling(kable(Tbl3_rep_n, caption = "Table 3R. Replicated No Weights. Age-based inequality in economic experience in percentages", col.names = c("Wave 1, Age 18-29", "Wave 1, Age 30-44", "Wave 1, Age 45-59", "Wave 1, 60+", "Wave 2, Age 18-29", "Wave 2, Age 30-44", "Wave 2, Age 45-59", "Wave 2, 60+")))
```
## Table 3O. Original

```{r Tbl3_orig}
colnames(Tbl3_orig) <- c("Variable", "Wave 1, Age 18-29", "Wave 1, Age 30-44", "Wave 1, Age 45-59", "Wave 1, 60+", "Wave 2, Age 18-29", "Wave 2, Age 30-44", "Wave 2, Age 45-59", "Wave 2, 60+")

Tbl3_orig$Variable <- c(NA, "stdliv_past5_fallen","stdliv_past5_risen", NA, "stdliv_next5_fallen","stdliv_next5_risen","noway_future_improve",NA, "mkt_econ_eval_pos","mkt_econ_eval_neg", "N")

Tbl3_orig <- subset(Tbl3_orig, !is.na(Variable))

kable_styling(kable(Tbl3_orig, caption = "Table 3O. Original. Age-based inequality in economic experience in percentages"))
```

### Table A.1

Just to check the sample

```{r tblA.1}
TblA1_repa <- df_orig_complete %>%
  subset(year == 1993) %>%
  group_by(cntry) %>%
  count(cntry) %>%
  ungroup()
  
TblA1_repb <- df_orig_complete %>%
  subset(year == 1994) %>%
  group_by(cntry) %>%
  count(cntry) %>%
  ungroup()
  
TblA1_repc <- df_orig_complete %>%
  subset(year == 1996) %>%
  group_by(cntry) %>%
  count(cntry) %>%
  ungroup()
  
TblA1_repd <- df_orig_complete %>%
  subset(year == 2007) %>%
  group_by(cntry) %>%
  count(cntry) %>%
  ungroup()
  
TblA1_rep <- left_join(TblA1_repd, TblA1_repa, by = "cntry")
TblA1_rep <- left_join(TblA1_rep, TblA1_repb, by = "cntry")
TblA1_rep <- left_join(TblA1_rep, TblA1_repc, by = "cntry")

TblA1_rep[,1:5] <- TblA1_rep[,c(1,3,4,5,2)]

colnames(TblA1_rep) <- c("country", "1993 N","1994 N","1996 N","2007 N")

TblA1r <- kable_styling(kable(TblA1_rep))

save_kable(TblA1r, file = "results/TblA1_rep.htm")

webshot("results/TblA1_rep.htm", file = "results/TblA1_rep.png")

kable_styling(kable(TblA1_rep, caption = "Table 1.A Replicated"))
```


## Save Final DF

```{r saveout}
# remove some unused variables to save filesize
df_orig_complete <- df_orig_complete %>%
  dplyr::select(stdliv_next5, stdliv_next5_di, stdliv_past5, stdliv_past5_di, cat_age, wave, female, education, EGP6, income, pensions, unemployed, car_owner, cntry, y1)

write.csv(df_orig_complete, here::here("data/df_orig_complete.csv"), row.names = F)
```

### Colophon
```{r colophon}
sessionInfo()
```

